---
title: "pulling_numbers"
author: "Anton Hung"
date: "2024-11-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) 
```

```{r}
library(data.table)
library(dplyr)

opioid <- read.csv("~/medicaid/medicaid_treatment_trends/output/results_df.csv")

alternate <- read.csv("~/medicaid/medicaid_treatment_trends/output/additional_outcomes_results.csv")
setDT(opioid)
opioid <- opioid[ ,-"X"]
setDT(alternate)
alternate <- alternate[ ,-"X"]

opioid[, num_opioid := number_of_beneficiaries*opioid_prop]
opioid[, total_mme := num_opioid*average_mme]
opioid[, total_days := num_opioid*average_days_covered]

# num physical therapy
setDT(alternate)
alternate[, num_physical_therapy := number_of_beneficiaries*physical_therapy_prop]
alternate[, num_chiropractic := number_of_beneficiaries*chiropractic_prop]
alternate[, num_acupuncture := number_of_beneficiaries*acupuncture_prop]
alternate[, num_nonopioid := number_of_beneficiaries*nonopioid_pain_prop]
alternate[, num_counselling := number_of_beneficiaries*counselling_prop]

```

# Opioid probability

## By pain
```{r}

sum(opioid[pain_or_disability=="chronic pain only", num_opioid] /
      sum(opioid[pain_or_disability=="chronic pain only", number_of_beneficiaries]))

sum(opioid[pain_or_disability=="disability only", num_opioid] /
      sum(opioid[pain_or_disability=="disability only", number_of_beneficiaries]))

sum(opioid[pain_or_disability=="disability and chronic pain", num_opioid] /
      sum(opioid[pain_or_disability=="disability and chronic pain", number_of_beneficiaries]))
```

## By year
```{r}
sum(opioid[year=="2016", num_opioid] /
      sum(opioid[year=="2016", number_of_beneficiaries]))
sum(opioid[year=="2017", num_opioid] /
      sum(opioid[year=="2017", number_of_beneficiaries]))
sum(opioid[year=="2018", num_opioid] /
      sum(opioid[year=="2018", number_of_beneficiaries]))
sum(opioid[year=="2019", num_opioid] /
      sum(opioid[year=="2019", number_of_beneficiaries]))
```


# MME
## by pain
```{r}


sum(opioid[pain_or_disability=="chronic pain only", total_mme] /
      sum(opioid[pain_or_disability=="chronic pain only", num_opioid]))

sum(opioid[pain_or_disability=="disability only", total_mme] /
      sum(opioid[pain_or_disability=="disability only", num_opioid]))

sum(opioid[pain_or_disability=="disability and chronic pain", total_mme] /
      sum(opioid[pain_or_disability=="disability and chronic pain", num_opioid]))
```

# Days covered
```{r}


sum(opioid[pain_or_disability=="chronic pain only", total_days] /
      sum(opioid[pain_or_disability=="chronic pain only", num_opioid]))

sum(opioid[pain_or_disability=="disability only", total_days] /
      sum(opioid[pain_or_disability=="disability only", num_opioid]))

sum(opioid[pain_or_disability=="disability and chronic pain", total_days] /
      sum(opioid[pain_or_disability=="disability and chronic pain", num_opioid]))
```

# Opioids among asians
```{r}
sum(opioid[race_ethnicity=="Asian, non-Hispanic" & pain_or_disability=="chronic pain only", num_opioid])/
  sum(opioid[race_ethnicity=="Asian, non-Hispanic" & pain_or_disability=="chronic pain only", number_of_beneficiaries])


sum(opioid[race_ethnicity=="Asian, non-Hispanic", num_opioid])/sum(opioid[race_ethnicity=="Asian, non-Hispanic", number_of_beneficiaries])

sum(opioid[race_ethnicity=="Asian, non-Hispanic", total_mme])/sum(opioid[race_ethnicity=="Asian, non-Hispanic", num_opioid])

sum(opioid[race_ethnicity=="Asian, non-Hispanic" & pain_or_disability=="chronic pain only", total_mme])/
  sum(opioid[race_ethnicity=="Asian, non-Hispanic" & pain_or_disability=="chronic pain only", num_opioid])

sum(opioid[race_ethnicity=="Asian, non-Hispanic", total_days])/sum(opioid[race_ethnicity=="Asian, non-Hispanic", num_opioid])

```

# opioids among whites
```{r}
sum(opioid[race_ethnicity=="White, non-Hispanic" & pain_or_disability=="chronic pain only", num_opioid])/
  sum(opioid[race_ethnicity=="White, non-Hispanic" & pain_or_disability=="chronic pain only", number_of_beneficiaries])

sum(opioid[race_ethnicity=="White, non-Hispanic", num_opioid])/
  sum(opioid[race_ethnicity=="White, non-Hispanic", number_of_beneficiaries])

sum(opioid[race_ethnicity=="White, non-Hispanic", total_mme])/sum(opioid[race_ethnicity=="White, non-Hispanic", num_opioid])

sum(opioid[race_ethnicity=="White, non-Hispanic" & pain_or_disability=="chronic pain only", total_mme])/
  sum(opioid[race_ethnicity=="White, non-Hispanic" & pain_or_disability=="chronic pain only", num_opioid])

sum(opioid[race_ethnicity=="White, non-Hispanic", total_days])/sum(opioid[race_ethnicity=="White, non-Hispanic", num_opioid])
```

# Opioids among blacks
```{r}
sum(opioid[race_ethnicity=="Black, non-Hispanic" & pain_or_disability=="chronic pain only", num_opioid])/
  sum(opioid[race_ethnicity=="Black, non-Hispanic" & pain_or_disability=="chronic pain only", number_of_beneficiaries])

sum(opioid[race_ethnicity=="Black, non-Hispanic", num_opioid])/
  sum(opioid[race_ethnicity=="Black, non-Hispanic", number_of_beneficiaries])

sum(opioid[race_ethnicity=="Black, non-Hispanic", total_mme])/sum(opioid[race_ethnicity=="Black, non-Hispanic", num_opioid])

sum(opioid[race_ethnicity=="Black, non-Hispanic", total_days])/sum(opioid[race_ethnicity=="Black, non-Hispanic", num_opioid])

```

# Opioids among AIAN
```{r}
sum(opioid[race_ethnicity=="AIAN_or_HPI" & pain_or_disability=="chronic pain only", num_opioid])/
  sum(opioid[race_ethnicity=="AIAN_or_HPI" & pain_or_disability=="chronic pain only", number_of_beneficiaries])

sum(opioid[race_ethnicity=="AIAN_or_HPI", num_opioid])/
  sum(opioid[race_ethnicity=="AIAN_or_HPI", number_of_beneficiaries])

sum(opioid[race_ethnicity=="AIAN_or_HPI", total_mme])/sum(opioid[race_ethnicity=="AIAN_or_HPI", num_opioid])

sum(opioid[race_ethnicity=="AIAN_or_HPI", total_days])/sum(opioid[race_ethnicity=="AIAN_or_HPI", num_opioid])

```


# Additional treatments 

```{r}
sum(alternate[, num_nonopioid])/
  sum(alternate[, number_of_beneficiaries])
sum(alternate[, num_counselling])/
  sum(alternate[, number_of_beneficiaries])

sum(alternate[pain_or_disability == "disability and chronic pain", num_nonopioid])/
  sum(alternate[pain_or_disability == "disability and chronic pain", number_of_beneficiaries])
sum(alternate[pain_or_disability == "disability and chronic pain", num_counselling])/
  sum(alternate[pain_or_disability == "disability and chronic pain", number_of_beneficiaries])

sum(alternate[pain_or_disability == "chronic pain only", num_physical_therapy])/
  sum(alternate[pain_or_disability == "chronic pain only", number_of_beneficiaries])

```


# Asians, pain and disability, PT and Acu
```{r}
sum(alternate[race_ethnicity=="Asian, non-Hispanic" & pain_or_disability=="disability and chronic pain", num_physical_therapy])/
  sum(alternate[race_ethnicity=="Asian, non-Hispanic" & pain_or_disability=="disability and chronic pain", number_of_beneficiaries])

sum(alternate[race_ethnicity=="Asian, non-Hispanic" & pain_or_disability=="disability and chronic pain", num_acupuncture])/
  sum(alternate[race_ethnicity=="Asian, non-Hispanic" & pain_or_disability=="disability and chronic pain", number_of_beneficiaries])


# Years 2016 vs 2019
sum(alternate[race_ethnicity=="Asian, non-Hispanic" & 
                pain_or_disability=="disability and chronic pain" &
                year=="2016", num_acupuncture])/
  sum(alternate[race_ethnicity=="Asian, non-Hispanic" & 
                pain_or_disability=="disability and chronic pain" &
                year=="2016", number_of_beneficiaries])
sum(alternate[race_ethnicity=="Asian, non-Hispanic" & 
                pain_or_disability=="disability and chronic pain" &
                year=="2019", num_acupuncture])/
  sum(alternate[race_ethnicity=="Asian, non-Hispanic" & 
                pain_or_disability=="disability and chronic pain" &
                year=="2019", number_of_beneficiaries])
```

Chiro, acu and PT for chronic pain only patients
```{r}
sum(alternate[pain_or_disability=="chronic pain only", num_chiropractic])/
  sum(alternate[pain_or_disability=="chronic pain only", number_of_beneficiaries])
sum(alternate[pain_or_disability=="chronic pain only", num_acupuncture])/
  sum(alternate[pain_or_disability=="chronic pain only", number_of_beneficiaries])
sum(alternate[pain_or_disability=="chronic pain only", num_physical_therapy])/
  sum(alternate[pain_or_disability=="chronic pain only", number_of_beneficiaries])
```

Chiro, acu and PT for pain + disability patients

```{r}
sum(alternate[pain_or_disability=="disability and chronic pain", num_chiropractic])/
  sum(alternate[pain_or_disability=="disability and chronic pain", number_of_beneficiaries])
sum(alternate[pain_or_disability=="disability and chronic pain", num_acupuncture])/
  sum(alternate[pain_or_disability=="disability and chronic pain", number_of_beneficiaries])
sum(alternate[pain_or_disability=="disability and chronic pain", num_physical_therapy])/
  sum(alternate[pain_or_disability=="disability and chronic pain", number_of_beneficiaries])
```

# Additional treatments: 2016 vs 2019
## counseling, pt, acupuncture, chiropractic 
```{r}
print("counseling")
sum(alternate[year == "2016", num_counselling])/
  sum(alternate[year == "2016", number_of_beneficiaries])*100
sum(alternate[year == "2019", num_counselling])/
  sum(alternate[year == "2019", number_of_beneficiaries])*100

print("pt")
sum(alternate[year == "2016", num_physical_therapy])/
  sum(alternate[year == "2016", number_of_beneficiaries])*100
sum(alternate[year == "2019", num_physical_therapy])/
  sum(alternate[year == "2019", number_of_beneficiaries])*100

print("acu")
sum(alternate[year == "2016", num_acupuncture])/
  sum(alternate[year == "2016", number_of_beneficiaries])*100
sum(alternate[year == "2019", num_acupuncture])/
  sum(alternate[year == "2019", number_of_beneficiaries])*100

print("chiro")
sum(alternate[year == "2016", num_chiropractic])/
  sum(alternate[year == "2016", number_of_beneficiaries])*100
sum(alternate[year == "2019", num_chiropractic])/
  sum(alternate[year == "2019", number_of_beneficiaries])*100
```


# TREATMENT Counseling + chiropractic
# AMONG White 
# VS Asian
# WITH CHRONIC PAIN
# IN THE YEAR 2019
```{r}
# see 02_06_subgroup_exploration.R
```



# TREATMENT Acupuncture + physical therapy
# AMONG ASIAN 
# VS WHITE
# WITH CHRONIC PAIN
# IN THE YEAR 2019
```{r}
# see 02_06_subgroup_exploration.R
```