---
title: "pulling_numbers"
author: "Anton Hung"
date: "2024-11-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) 
```

```{r}
library(data.table)
library(dplyr)
result_dir <- "~/medicaid/medicaid-treatment-trends/output"

# opioid <- read.csv("/Users/anton/Library/CloudStorage/OneDrive-ColumbiaUniversityIrvingMedicalCenter/2025/magda/new_results/results_250513.csv")
opioid <- read.csv(file.path(result_dir, "results_250513.csv"))

# alternate <- read.csv("/Users/anton/Library/CloudStorage/OneDrive-ColumbiaUniversityIrvingMedicalCenter/2025/magda/new_results/additional_outcomes_results_25_05_13.csv")
alternate <- read.csv(file.path(result_dir, "additional_outcomes_results_25_05_13.csv"))

setDT(opioid)
opioid <- opioid[ ,-"X"]
setDT(alternate)
alternate <- alternate[ ,-"X"] |>
  mutate(n_acupuncture_outpatient = as.numeric(n_acupuncture_outpatient))


# opioid[, n_opioid_outpatient := number_of_beneficiaries*opioid_prop]
opioid[, total_mme := n_opioid_outpatient*average_mme_outpatient]
opioid[, total_days := n_opioid_outpatient*average_days_covered_outpatient]

# num physical therapy
# alternate[, num_nonopioid := number_of_beneficiaries*counselling_prop]
# alternate[, num_physical_therapy := number_of_beneficiaries*physical_therapy_prop]
# alternate[, num_chiropractic := number_of_beneficiaries*chiropractic_prop]
# alternate[, num_acupuncture := number_of_beneficiaries*acupuncture_prop]
# alternate[, num_nonopioid := number_of_beneficiaries*nonopioid_pain_prop]
# alternate[, num_counselling := number_of_beneficiaries*counselling_prop]

```

# Any treatment
```{r}
sum(alternate[pain_or_disability=="chronic pain only", n_any_treatment] /
      sum(alternate[pain_or_disability=="chronic pain only", number_of_beneficiaries]))

sum(alternate[pain_or_disability=="disability and chronic pain", n_any_treatment] /
      sum(alternate[pain_or_disability=="disability and chronic pain", number_of_beneficiaries]))

```
# Among white with chronic pain, asian with chronic pain
```{r}
print("White")
sum(alternate[pain_or_disability=="chronic pain only" &
                race_ethnicity=="White, non-Hispanic", n_any_treatment] /
      sum(alternate[pain_or_disability=="chronic pain only" &
                race_ethnicity=="White, non-Hispanic", number_of_beneficiaries]))

print("Asian")
sum(alternate[pain_or_disability=="chronic pain only" &
                race_ethnicity=="Asian, non-Hispanic", n_any_treatment] /
      sum(alternate[pain_or_disability=="chronic pain only" &
                race_ethnicity=="Asian, non-Hispanic", number_of_beneficiaries]))

```


# Pharmacologic
## all opioid vs non-opioid
```{r}
sum(opioid[, n_opioid_outpatient] /
      sum(opioid[, number_of_beneficiaries]))
sum(opioid[, n_nonopioid_outpatient] /
      sum(opioid[, number_of_beneficiaries]))

```

# Opioid probability

## By pain
```{r}

sum(opioid[pain_or_disability=="chronic pain only", n_opioid_outpatient] /
      sum(opioid[pain_or_disability=="chronic pain only", number_of_beneficiaries]))

# sum(opioid[pain_or_disability=="disability only", n_opioid_outpatient] /
#       sum(opioid[pain_or_disability=="disability only", number_of_beneficiaries]))

sum(opioid[pain_or_disability=="disability and chronic pain", n_opioid_outpatient] /
      sum(opioid[pain_or_disability=="disability and chronic pain", number_of_beneficiaries]))
```

## By year
```{r}
sum(opioid[year=="2016", n_opioid_outpatient] /
      sum(opioid[year=="2016", number_of_beneficiaries]))
sum(opioid[year=="2017", n_opioid_outpatient] /
      sum(opioid[year=="2017", number_of_beneficiaries]))
sum(opioid[year=="2018", n_opioid_outpatient] /
      sum(opioid[year=="2018", number_of_beneficiaries]))
sum(opioid[year=="2019", n_opioid_outpatient] /
      sum(opioid[year=="2019", number_of_beneficiaries]))
```


# MME
## by pain
```{r}


sum(opioid[pain_or_disability=="chronic pain only", total_mme] /
      sum(opioid[pain_or_disability=="chronic pain only", n_opioid_outpatient]))

# sum(opioid[pain_or_disability=="disability only", total_mme] /
#       sum(opioid[pain_or_disability=="disability only", n_opioid_outpatient]))

sum(opioid[pain_or_disability=="disability and chronic pain", total_mme] /
      sum(opioid[pain_or_disability=="disability and chronic pain", n_opioid_outpatient]))
```

# Days covered
```{r}


sum(opioid[pain_or_disability=="chronic pain only", total_days] /
      sum(opioid[pain_or_disability=="chronic pain only", n_opioid_outpatient]))

# sum(opioid[pain_or_disability=="disability only", total_days] /
#       sum(opioid[pain_or_disability=="disability only", n_opioid_outpatient]))

sum(opioid[pain_or_disability=="disability and chronic pain", total_days] /
      sum(opioid[pain_or_disability=="disability and chronic pain", n_opioid_outpatient]))
```

# Opioids among asians
```{r}
sum(opioid[race_ethnicity=="Asian, non-Hispanic" & pain_or_disability=="chronic pain only", n_opioid_outpatient])/
  sum(opioid[race_ethnicity=="Asian, non-Hispanic" & pain_or_disability=="chronic pain only", number_of_beneficiaries])


sum(opioid[race_ethnicity=="Asian, non-Hispanic", n_opioid_outpatient])/sum(opioid[race_ethnicity=="Asian, non-Hispanic", number_of_beneficiaries])

sum(opioid[race_ethnicity=="Asian, non-Hispanic", total_mme])/sum(opioid[race_ethnicity=="Asian, non-Hispanic", n_opioid_outpatient])

sum(opioid[race_ethnicity=="Asian, non-Hispanic" & pain_or_disability=="chronic pain only", total_mme])/
  sum(opioid[race_ethnicity=="Asian, non-Hispanic" & pain_or_disability=="chronic pain only", n_opioid_outpatient])

sum(opioid[race_ethnicity=="Asian, non-Hispanic", total_days])/sum(opioid[race_ethnicity=="Asian, non-Hispanic", n_opioid_outpatient])

```

# opioids among whites
```{r}
sum(opioid[race_ethnicity=="White, non-Hispanic" & pain_or_disability=="chronic pain only", n_opioid_outpatient])/
  sum(opioid[race_ethnicity=="White, non-Hispanic" & pain_or_disability=="chronic pain only", number_of_beneficiaries])

sum(opioid[race_ethnicity=="White, non-Hispanic", n_opioid_outpatient])/
  sum(opioid[race_ethnicity=="White, non-Hispanic", number_of_beneficiaries])

sum(opioid[race_ethnicity=="White, non-Hispanic", total_mme])/sum(opioid[race_ethnicity=="White, non-Hispanic", n_opioid_outpatient])

sum(opioid[race_ethnicity=="White, non-Hispanic" & pain_or_disability=="chronic pain only", total_mme])/
  sum(opioid[race_ethnicity=="White, non-Hispanic" & pain_or_disability=="chronic pain only", n_opioid_outpatient])

sum(opioid[race_ethnicity=="White, non-Hispanic", total_days])/sum(opioid[race_ethnicity=="White, non-Hispanic", n_opioid_outpatient])
```

# Opioids among blacks
```{r}
sum(opioid[race_ethnicity=="Black, non-Hispanic" & pain_or_disability=="chronic pain only", n_opioid_outpatient])/
  sum(opioid[race_ethnicity=="Black, non-Hispanic" & pain_or_disability=="chronic pain only", number_of_beneficiaries])

sum(opioid[race_ethnicity=="Black, non-Hispanic", n_opioid_outpatient])/
  sum(opioid[race_ethnicity=="Black, non-Hispanic", number_of_beneficiaries])

sum(opioid[race_ethnicity=="Black, non-Hispanic", total_mme])/sum(opioid[race_ethnicity=="Black, non-Hispanic", n_opioid_outpatient])

sum(opioid[race_ethnicity=="Black, non-Hispanic", total_days])/sum(opioid[race_ethnicity=="Black, non-Hispanic", n_opioid_outpatient])

```

# Opioids among AIAN
```{r}
sum(opioid[race_ethnicity=="AIAN_or_HPI" & pain_or_disability=="chronic pain only", n_opioid_outpatient])/
  sum(opioid[race_ethnicity=="AIAN_or_HPI" & pain_or_disability=="chronic pain only", number_of_beneficiaries])

sum(opioid[race_ethnicity=="AIAN_or_HPI", n_opioid_outpatient])/
  sum(opioid[race_ethnicity=="AIAN_or_HPI", number_of_beneficiaries])

sum(opioid[race_ethnicity=="AIAN_or_HPI", total_mme])/sum(opioid[race_ethnicity=="AIAN_or_HPI", n_opioid_outpatient])

sum(opioid[race_ethnicity=="AIAN_or_HPI", total_days])/sum(opioid[race_ethnicity=="AIAN_or_HPI", n_opioid_outpatient])

```

# Opioids among Hispanic
```{r}
print("MME")
sum(opioid[race_ethnicity=="Hispanic, all races", total_mme])/sum(opioid[race_ethnicity=="Hispanic, all races", n_opioid_outpatient])

print("Days covered")
sum(opioid[race_ethnicity=="Hispanic, all races", total_days])/sum(opioid[race_ethnicity=="Hispanic, all races", n_opioid_outpatient])

```

# Additional treatments: 2016 vs 2019
## counseling, pt, acupuncture, chiropractic, intervention 
```{r}
print("counseling")
sum(alternate[year == "2016", n_counseling_outpatient])/
  sum(alternate[year == "2016", number_of_beneficiaries])*100
sum(alternate[year == "2019", n_counseling_outpatient])/
  sum(alternate[year == "2019", number_of_beneficiaries])*100

print("pt")
sum(alternate[year == "2016", n_physical_therapy_outpatient])/
  sum(alternate[year == "2016", number_of_beneficiaries])*100
sum(alternate[year == "2019", n_physical_therapy_outpatient])/
  sum(alternate[year == "2019", number_of_beneficiaries])*100

print("acu")
sum(alternate[year == "2016", n_acupuncture_outpatient], na.rm=T)/
  sum(alternate[year == "2016", number_of_beneficiaries])*100
sum(alternate[year == "2019", n_acupuncture_outpatient], na.rm=T)/
  sum(alternate[year == "2019", number_of_beneficiaries])*100

print("chiro")
sum(alternate[year == "2016", n_chiropractic_outpatient])/
  sum(alternate[year == "2016", number_of_beneficiaries])*100
sum(alternate[year == "2019", n_chiropractic_outpatient])/
  sum(alternate[year == "2019", number_of_beneficiaries])*100

print("intervention")
sum(alternate[year == "2016", n_intervention_outpatient])/
  sum(alternate[year == "2016", number_of_beneficiaries])*100
sum(alternate[year == "2019", n_intervention_outpatient])/
  sum(alternate[year == "2019", number_of_beneficiaries])*100
```


# Asians, pain and disability, PT and Acu
```{r}
sum(alternate[race_ethnicity=="Asian, non-Hispanic" & pain_or_disability=="disability and chronic pain", n_physical_therapy_outpatient])/
  sum(alternate[race_ethnicity=="Asian, non-Hispanic" & pain_or_disability=="disability and chronic pain", number_of_beneficiaries])

sum(alternate[race_ethnicity=="Asian, non-Hispanic" & pain_or_disability=="disability and chronic pain", n_acupuncture_outpatient])/
  sum(alternate[race_ethnicity=="Asian, non-Hispanic" & pain_or_disability=="disability and chronic pain", number_of_beneficiaries])

sum(alternate[race_ethnicity=="Asian, non-Hispanic" & pain_or_disability=="chronic pain only", n_physical_therapy_outpatient])/
  sum(alternate[race_ethnicity=="Asian, non-Hispanic" & pain_or_disability=="chronic pain only", number_of_beneficiaries])

sum(alternate[race_ethnicity=="Asian, non-Hispanic" & pain_or_disability=="chronic pain only", n_acupuncture_outpatient])/
  sum(alternate[race_ethnicity=="Asian, non-Hispanic" & pain_or_disability=="chronic pain only", number_of_beneficiaries])

# Years 2016 vs 2019
sum(alternate[race_ethnicity=="Asian, non-Hispanic" & 
                pain_or_disability=="chronic pain only" &
                year=="2016", n_acupuncture_outpatient])/
  sum(alternate[race_ethnicity=="Asian, non-Hispanic" & 
                pain_or_disability=="chronic pain only" &
                year=="2016", number_of_beneficiaries])
sum(alternate[race_ethnicity=="Asian, non-Hispanic" & 
                pain_or_disability=="chronic pain only" &
                year=="2019", n_acupuncture_outpatient])/
  sum(alternate[race_ethnicity=="Asian, non-Hispanic" & 
                pain_or_disability=="chronic pain only" &
                year=="2019", number_of_beneficiaries])

sum(alternate[race_ethnicity=="Asian, non-Hispanic" & 
                pain_or_disability=="disability and chronic pain" &
                year=="2016", n_acupuncture_outpatient])/
  sum(alternate[race_ethnicity=="Asian, non-Hispanic" & 
                pain_or_disability=="disability and chronic pain" &
                year=="2016", number_of_beneficiaries])
sum(alternate[race_ethnicity=="Asian, non-Hispanic" & 
                pain_or_disability=="disability and chronic pain" &
                year=="2019", n_acupuncture_outpatient])/
  sum(alternate[race_ethnicity=="Asian, non-Hispanic" & 
                pain_or_disability=="disability and chronic pain" &
                year=="2019", number_of_beneficiaries])
```

Chiro, Acupuncture and PT
```{r}
print("chronic pain")
sum(alternate[pain_or_disability=="chronic pain only", n_chiropractic_outpatient])/
  sum(alternate[pain_or_disability=="chronic pain only", number_of_beneficiaries])
sum(alternate[pain_or_disability=="chronic pain only", n_acupuncture_outpatient])/
  sum(alternate[pain_or_disability=="chronic pain only", number_of_beneficiaries])
sum(alternate[pain_or_disability=="chronic pain only", n_physical_therapy_outpatient])/
  sum(alternate[pain_or_disability=="chronic pain only", number_of_beneficiaries])

print("pain and disability")
sum(alternate[pain_or_disability=="disability and chronic pain", n_chiropractic_outpatient])/
  sum(alternate[pain_or_disability=="disability and chronic pain", number_of_beneficiaries])
sum(alternate[pain_or_disability=="disability and chronic pain", n_acupuncture_outpatient], na.rm=T)/
  sum(alternate[pain_or_disability=="disability and chronic pain", number_of_beneficiaries])
sum(alternate[pain_or_disability=="disability and chronic pain", n_physical_therapy_outpatient])/
  sum(alternate[pain_or_disability=="disability and chronic pain", number_of_beneficiaries])
```

Counseling, Interventions
```{r}
print("chronic pain")
sum(alternate[pain_or_disability=="chronic pain only", n_counseling_outpatient])/
  sum(alternate[pain_or_disability=="chronic pain only", number_of_beneficiaries])
sum(alternate[pain_or_disability=="chronic pain only", n_intervention_outpatient])/
  sum(alternate[pain_or_disability=="chronic pain only", number_of_beneficiaries])

print("pain and disability")
sum(alternate[pain_or_disability=="disability and chronic pain", n_counseling_outpatient])/
  sum(alternate[pain_or_disability=="disability and chronic pain", number_of_beneficiaries])
sum(alternate[pain_or_disability=="disability and chronic pain", n_intervention_outpatient], na.rm=T)/
  sum(alternate[pain_or_disability=="disability and chronic pain", number_of_beneficiaries])

```

# White, chronic pain alone
# Counseling, Chiropractic, Interventions
```{r}
print("counseling")
sum(alternate[pain_or_disability=="chronic pain only" &
                race_ethnicity=="White, non-Hispanic", n_counseling_outpatient])/
  sum(alternate[pain_or_disability=="chronic pain only" &
                race_ethnicity=="White, non-Hispanic", number_of_beneficiaries])
print("chiropractic")
sum(alternate[pain_or_disability=="chronic pain only" &
                race_ethnicity=="White, non-Hispanic", n_chiropractic_outpatient])/
  sum(alternate[pain_or_disability=="chronic pain only" &
                race_ethnicity=="White, non-Hispanic", number_of_beneficiaries])
print("interventions")
sum(alternate[pain_or_disability=="chronic pain only" &
                race_ethnicity=="White, non-Hispanic", n_intervention_outpatient])/
  sum(alternate[pain_or_disability=="chronic pain only" &
                race_ethnicity=="White, non-Hispanic", number_of_beneficiaries])
```

# White, disability and pain
# Counseling, Chiropractic, Interventions
```{r}
print("counseling")
sum(alternate[pain_or_disability=="disability and chronic pain" &
                race_ethnicity=="White, non-Hispanic", n_counseling_outpatient])/
  sum(alternate[pain_or_disability=="disability and chronic pain" &
                race_ethnicity=="White, non-Hispanic", number_of_beneficiaries])
print("chiropractic")
sum(alternate[pain_or_disability=="disability and chronic pain" &
                race_ethnicity=="White, non-Hispanic", n_chiropractic_outpatient])/
  sum(alternate[pain_or_disability=="disability and chronic pain" &
                race_ethnicity=="White, non-Hispanic", number_of_beneficiaries])
print("interventions")
sum(alternate[pain_or_disability=="disability and chronic pain" &
                race_ethnicity=="White, non-Hispanic", n_intervention_outpatient])/
  sum(alternate[pain_or_disability=="disability and chronic pain" &
                race_ethnicity=="White, non-Hispanic", number_of_beneficiaries])
```

# Multiracial, disability and pain
# Counseling, Chiropractic, Interventions
```{r}
print("Counseling")
sum(alternate[pain_or_disability=="disability and chronic pain" &
                race_ethnicity=="Multiracial or Missing", n_counseling_outpatient])/
  sum(alternate[pain_or_disability=="disability and chronic pain" &
                race_ethnicity=="Multiracial or Missing", number_of_beneficiaries])
print("chiropractic")
sum(alternate[pain_or_disability=="disability and chronic pain" &
                race_ethnicity=="Multiracial or Missing", n_chiropractic_outpatient])/
  sum(alternate[pain_or_disability=="disability and chronic pain" &
                race_ethnicity=="Multiracial or Missing", number_of_beneficiaries])
print("interventions")
sum(alternate[pain_or_disability=="disability and chronic pain" &
                race_ethnicity=="Multiracial or Missing", n_intervention_outpatient])/
  sum(alternate[pain_or_disability=="disability and chronic pain" &
                race_ethnicity=="Multiracial or Missing", number_of_beneficiaries])
```


# TREATMENT Counseling + chiropractic
# AMONG White 
# VS Asian
# WITH CHRONIC PAIN
# IN THE YEAR 2019
```{r}
# see 02_06_subgroup_exploration.R

print("White, chiro")
sum(alternate[year == "2019" &
                race_ethnicity =="White, non-Hispanic" &
                pain_or_disability=="chronic pain only", n_chiropractic_outpatient])/
  sum(alternate[year == "2019" &
                race_ethnicity =="White, non-Hispanic" &
                pain_or_disability=="chronic pain only", number_of_beneficiaries])*100
print("White, couns")
sum(alternate[year == "2019" &
                race_ethnicity =="White, non-Hispanic" &
                pain_or_disability=="chronic pain only", n_counseling_outpatient])/
  sum(alternate[year == "2019" &
                race_ethnicity =="White, non-Hispanic" &
                pain_or_disability=="chronic pain only", number_of_beneficiaries])*100
print("White, intervention")
sum(alternate[year == "2019" &
                race_ethnicity =="White, non-Hispanic" &
                pain_or_disability=="chronic pain only", n_intervention_outpatient])/
  sum(alternate[year == "2019" &
                race_ethnicity =="White, non-Hispanic" &
                pain_or_disability=="chronic pain only", number_of_beneficiaries])*100
print("Asian, chiro")
sum(alternate[year == "2019" &
                race_ethnicity =="Asian, non-Hispanic" &
                pain_or_disability=="chronic pain only", n_chiropractic_outpatient])/
  sum(alternate[year == "2019" &
                race_ethnicity =="Asian, non-Hispanic" &
                pain_or_disability=="chronic pain only", number_of_beneficiaries])*100
print("Asian, couns")
sum(alternate[year == "2019" &
                race_ethnicity =="Asian, non-Hispanic" &
                pain_or_disability=="chronic pain only", n_counseling_outpatient])/
  sum(alternate[year == "2019" &
                race_ethnicity =="Asian, non-Hispanic" &
                pain_or_disability=="chronic pain only", number_of_beneficiaries])*100
print("Asian, intervention")
sum(alternate[year == "2019" &
                race_ethnicity =="Asian, non-Hispanic" &
                pain_or_disability=="chronic pain only", n_intervention_outpatient])/
  sum(alternate[year == "2019" &
                race_ethnicity =="Asian, non-Hispanic" &
                pain_or_disability=="chronic pain only", number_of_beneficiaries])*100
print("White, acup")
sum(alternate[year == "2019" &
                race_ethnicity =="White, non-Hispanic" &
                pain_or_disability=="chronic pain only", n_acupuncture_outpatient])/
  sum(alternate[year == "2019" &
                race_ethnicity =="White, non-Hispanic" &
                pain_or_disability=="chronic pain only", number_of_beneficiaries])*100
print("White, pt")
sum(alternate[year == "2019" &
                race_ethnicity =="White, non-Hispanic" &
                pain_or_disability=="chronic pain only", n_physical_therapy_outpatient])/
  sum(alternate[year == "2019" &
                race_ethnicity =="White, non-Hispanic" &
                pain_or_disability=="chronic pain only", number_of_beneficiaries])*100
print("Asian, acup")
sum(alternate[year == "2019" &
                race_ethnicity =="Asian, non-Hispanic" &
                pain_or_disability=="chronic pain only", n_acupuncture_outpatient])/
  sum(alternate[year == "2019" &
                race_ethnicity =="Asian, non-Hispanic" &
                pain_or_disability=="chronic pain only", number_of_beneficiaries])*100
print("Asian, pt")
sum(alternate[year == "2019" &
                race_ethnicity =="Asian, non-Hispanic" &
                pain_or_disability=="chronic pain only", n_physical_therapy_outpatient])/
  sum(alternate[year == "2019" &
                race_ethnicity =="Asian, non-Hispanic" &
                pain_or_disability=="chronic pain only", number_of_beneficiaries])*100
```



# TREATMENT Acupuncture + physical therapy
# AMONG ASIAN 
# VS WHITE
# WITH CHRONIC PAIN
# IN THE YEAR 2019
```{r}
# see 02_06_subgroup_exploration.R
```


# Counseling and non-opioid pharma
# Aong beneficiaries with pain + disability vs pain alone
```{r}
### pain + disability
# counseling
sum(alternate[pain_or_disability=="disability and chronic pain", n_counseling_outpatient])/
  sum(alternate[pain_or_disability=="disability and chronic pain", number_of_beneficiaries])
# nonopioid pharma
sum(opioid[pain_or_disability=="disability and chronic pain", n_nonopioid_outpatient])/
  sum(opioid[pain_or_disability=="disability and chronic pain", number_of_beneficiaries])

### pain alone
sum(alternate[pain_or_disability=="chronic pain only", n_counseling_outpatient])/
  sum(alternate[pain_or_disability=="chronic pain only", number_of_beneficiaries])
# nonopioid pharma
sum(opioid[pain_or_disability=="chronic pain only", n_nonopioid_outpatient])/
  sum(opioid[pain_or_disability=="chronic pain only", number_of_beneficiaries])
```





